\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-06-22 eg6-letters/NIM/EG6_RTPC_NIM.tex
NIM:  add resolution and elastic scattering paragraphs

The previous discussions have shown that the two modules of the RTPC have 
slightly different behavior and give also different particle yield. However, 
this should not necessarily be linked to a different performance of the RTPC. 
Indeed, there is a complicated convolution of CLAS and the RTPC acceptances 
that can significantly affect the yields and the kinematic of the detected 
particles. To clarify this question, we measured the 
efficiency of the RTPC using the elastic scattering on $^4$He, by comparing the 
inclusive yield based only on electron detection to the exclusive elastic 
yield where the Helium recoil is also detected. We present in 
Fig.~\ref{fig:rtpc_eff} the results for the two modules 
of the detector. We observe that the left and 
the right modules have similar efficiencies except near the target windows.

>>>>

The primary data sample used for calibration and performance assessment of the
RTPC was elastic scattering with a 1.2 GeV electron beam.  The electron's momentum
and direction is measured with CLAS, which uniquely determines the expected recoiling
$^4$He kinematics.  Matching requirements between reconstructed and expected $z$-vertex
and direction of the RTPC track provides a clean selection of elastically-scattered
$^4$He, showin in Fig.~\ref{fig:w}.

\begin{figure}[tb]\centering
  \includegraphics[width=8cm]{fig/fit_W_distribution_l.png}
  \caption{The $W$ distribution calculated from electron kinematics before (``inclusive'')
  and after (``exclusive'') requiring a matching track in the RTPC.\label{fig:w}}
\end{figure}

\subsection{Resolution}
Elastic scattering was used to estimate the tracking resolution of the RTPC 
based on the residual between the expected and measured $^4$He tracks.  The RTPC
resolutions, after removing contributions from the electron, are shown in Table
~\ref{tab:reso}, and are very similar for the two halves of the RTPC.  Note, the
$\theta$- and $z$-resolutions are highly correlated.
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|cccc|}
  \hline
& $\sigma_{z}$ &  $\sigma_{\theta}$ & $\sigma_{\phi}$ & $\sigma_{p}$\\
\hline
Left &  6.0 mm & 3.8$^{\circ}$ & 1.9$^{\circ}$ & 9$\%$ \\
Right & 7.4 mm & 4.0$^{\circ}$ & 1.9$^{\circ}$ & 8$\%$\\
\hline
\end{tabular}
\caption{The resolutions of the two modules of the RTPC for $z$-vertex, polar and azimuthal angles, and momentum.}
\label{tab:reso}
\end{center}
\end{table}

\subsection{Efficiency}
The previous discussions have shown that the two modules of the RTPC have 
slightly different behavior and give also different particle yield. However, 
this should not necessarily be linked to a different performance of the RTPC. 
Indeed, there is a complicated convolution of CLAS and the RTPC acceptances 
that can significantly affect the yields and kinematics of the detected 
particles. To clarify this question, we measured the efficiency of the RTPC
using elastic scattering on $^4$He by comparing the inclusive yield, based
only on electron detection, to the exclusive elastic yield, where the Helium
recoil is also detected (see Fig.~\ref{fig:w}). We present in Fig.~\ref{fig:rtpc_eff}
the results for the two modules of the detector. We observe that the left and 
the right modules have similar efficiencies except near the target windows.
